---
title: "Homework 5"
author: "Joel Pantuso"
date: "4/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(GGally)
library(ggfortify)
library(emmeans)
library(lindia)
```

## Problem 1

For this analysis, we will focus on predicting the List.price based on the year of construction, the number of bedrooms and baths, the square footage of the home, the size of the lot, and whether or not the home has a fireplace.

### Part 1

Read in the data and remove any columns from the data set that are not needed for modeling.
```{r}
houses <- read.csv(file="MountPleasantRealEstateData.csv")
houses <- houses %>%
  subset(select = c(List.Price, Year.Built, Bedrooms, Baths, Square.Footage, Acreage, Fireplace))
head(houses)
```

### Part 2

Construct a scatterplot matrix of all the relevant variables. Describe any discernible patterns you see.

```{r, message=FALSE}
ggpairs(houses, columns=1:7)
```

**I notice that square footage seems to be the most strongly correlated with list price. Unsurprisingly, bedrooms and bathrooms are also moderately correlated, as are both of these variables with square footage. Finally, the data for list price, acreage, and (to a lesser extent) square footage seems to be skewed right.**

### Part 3

Fit the multiple regression modeling List.Price as a function of Bedrooms, Baths, Square.Footage, Year.Built, Acreage and Fireplace and construct residual diagnostic plots of the fit. Do there appear to be any assumption violations?

```{r}
houses.fit <- lm(List.Price ~ Bedrooms + Baths + Square.Footage + Year.Built + Acreage + Fireplace, data=houses)
autoplot(houses.fit)
```

**There appear to be serious normality violations (plenty of high residual observations in the Normal Q-Q plot), as well as considerable constant variance violations (non-symmetric spread of residuals in residuals vs fitted plot). There may also be influential points present (109, 107, 144) based on the residuals vs leverage plot.**

### Part 4

Use a Box-Cox plot to determine whether a transformation of the response variable could improve the model fit. If the plot suggests a transformation, fit the model with that transformation and comment on the residual diagnostic plots for the revised model. Does it make sense to do a transformation?

```{r}
gg_boxcox(houses.fit)
```

**Since the boxcox plot includes 0 in the range of suggested transformations, it makes sense to perform a log transformation on the response variable.**

```{r}
houses.logfit <- lm(log10(List.Price) ~ Bedrooms + Baths + Square.Footage + Year.Built + Acreage + Fireplace, data=houses)
autoplot(houses.logfit)
```

**The Normal Q-Q and residuals vs fitted plots both look a lot better (variance looks constant now, normality looks fine (just influential points skewing it now)). Influential points still appear to be present, but this is not as severe of an issue.**

### Part 5

Is the model in part 4 significant in predicting selling price? Cite the test statistic, degrees of freedom, p-value and cite your conclusion in context.

```{r}
summary(houses.logfit)
```

**With an F-stat of 246 on 6 and 238 degrees of freedom (p=<2.2e-16), we can confidently say that the model from part 4 is significant in predicting the true mean selling price.**

### Part 6

What are the significant predictors in the model fitted in part 4? If there are insignificant predictors, discuss why they are not significant even though they may have a clear association with the response variable based on the scatterplot in part 2.

**Every predictor (including FireplaceYes & FireplaceNo (the intercept)) except Bedrooms is significant for the model from part 4. The reason that Bedrooms is not significant is likely because its effect is already accounted for by Baths, as these two variables were moderately correlated in the scatterplot from part 2.**

### Part 7

Use the model fitted in part 4 to construct a 95% confidence interval for the true mean selling price of all 4000 ft2 homes built in 2000 on 1 acre lots that have four bedrooms, three baths and a fireplace.  Be sure to completely interpret the interval in meaningful context.

```{r}
predict(houses.logfit, newdata=data.frame(Square.Footage=4000, Year.Built=2000, Acreage=1, Bedrooms=4, Baths=3, Fireplace="Yes"), int="conf")
```

**We can predict, with 95% confidence, that the true mean logarithm of the selling price for all 4000 ft2 homes built in 2000 on 1 acre lots that have four bedrooms, three baths and a fireplace is between 5.89 and 5.94.**

## Problem 2

```{r}
rat <- read.csv(file="ratlethargy.csv") %>%
  mutate(dose = as.factor(dose))
```


### Part 1

Perform a one-way ANOVA (see Module 3) to study the effect of dose alone on the rest time response variable. This analysis will completely ignore the effect of the age of the rats.  Check the underlying assumptions for the one-way ANOVA.  Do you see any problems?  If so, what is/are the problem(s)?

```{r}
rat.anova <- aov(resttime ~ dose, data=rat)
autoplot(rat.anova)
```

**There are indeed problems with the one-way ANOVA assumptions for this model. The residuals vs. fitted and scale-location plots both show extreme fanning, indicating a violation of the constant variance assumption.**

### Part 2

Now, we will consider performing an ANCOVA instead, studying the effect of dosage but controlling for the effect of age of the rats.  Start by creating a scatterplot of rest time vs. age by dosage level, with regression lines overlaid for each dosage group (use coloring to distinguish dosage groups and the lines). Comment clearly, in context, on what you see.  Does it appear that dosage of the drug has an effect on the induction of lethargy?

```{r, message=FALSE}
ggplot(rat, aes(x=age, y=resttime, color=dose)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, aes(color=dose)) +
  theme_minimal()
```

**Seeing how none of the trend lines are parallel to each other, it would appear that the dosage of the drug has an effect on lethargy. Based on what I can tell, it seems that the higher the dose, the longer the rest time, and that rest time increases with age.**

### Part 3

Fit a full interaction ANCOVA model to these data, and then check the assumptions.  Do there appear to be any assumption problems now?  Based on this, what appeared to be the source of the assumption problems in the one-way ANOVA in part 1?

```{r}
rat.ancova <- lm(resttime ~ dose + age + dose:age, data=rat)
autoplot(rat.ancova)
```

**The constant variance violation has gone away! The residuals are roughly symmetric about 0 in the residuals vs fitted plot now, indicating that we're good to go. Based on this, the source of this issue for the one-way ANOVA was that we did not account for the effect of rat age on rest time.**

### Part 4

Run a reduced F-test to see if the interaction between age and dosage level is significant.  Report your results in context, providing F-statistic value, degrees of freedom, and p-value. (Note that this test essentially asks if the effect of dosage level on induction of lethargy depends on the age of the rat).

```{r}
summary(rat.ancova)
```

**With an F-stat of 510.6 on 7 and 52 degrees of freedom (p=<2.2e-16), There is sufficient evidence to conclude that the effect of dosage level on induction of lethargy does depend on the age of the rat, with induction of lethargy increasing as dose increases.**

### Part 5

Based on your answer to part 4, are you able to make a general statement about the effect of dosage on the induction of lethargy?  Why or why not?  What is the complicating issue in trying to make such a broad statement?

**I cannot make a general statement about the effect of dosage on the induction of lethargy because I only have data for doses from 0-30. If I were to make that broad statement, that would imply that this effect is consistent across higher dosages as well, when in reality it may not be. We just can't tell without having any corresponding data.**

### Part 6

Use the appropriate model to make two different 95% prediction intervals for the rest time of an individual rat under each of the following conditions, being sure to interpret and comment on the collective findings:
-For a 11-month old rat that receives a 10mg dose
-For a 11-month old rat that receives a 30 mg dose

```{r}
predict(rat.ancova, newdata=data.frame(age=11, dose="10"), int="pred")
predict(rat.ancova, newdata=data.frame(age=11, dose="30"), int="pred")
```

**We can predict, with 95% confidence, that the rest time for a single randomly selected 11 month old rat that receives a 10 mg dose will be between 91.11 and 123.90 minutes. Similarly, we can predict, with 95% confidence, that the rest time for a single randomly selected 11 month old rat that receives a 30 mg dose will be between 183.04 and 215.75 minutes.**
