---
title: "M. iowae Width Measurements"
author: "Joel Pantuso"
date: "3/28/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(GGally)
library(ggfortify)
tinytex::install_tinytex()
```

## Introduction

For this experiment, Mycoplasma iowae cells were grown at 1% O2 (experimental) and 21% O2 (control) at 37°C for 48 hours. 3 replicates were performed for the experimental condition, whereas 2 replicates were performed for the control condition. Samples were then collected, prepped for SEM imaging, and imaged by Nathan.

The first control replicate was imaged without a scale bar and contained contaminant, thus it was omitted from data analysis. All available 5kx images from the remaining experiments were used for analysis, with the exception of the first experimental replicate which had none (and thus 3kx images were used, with measurements scaled accordingly). 

Cell widths were measured using Digimizer. 100 measurements were taken from each experimental image (n = 1100), and 367, 367, and 366 images were taken from each control image sequentially (n = 1100). Each image's measurements are represented below as low# or reg# (low = 1% O2, reg = 21% O2). All 1% O2 measurements were compiled into one dataset (alllow) for summary statistics, and the same was done with all 21% O2 measurements (allreg). Finally, all measurements were compiled into a giant dataset, represented below as oxygen, for EDA and ANOVA.


```{r read, message=FALSE}
low1 <- read_csv("rep1_3kx_1.csv")
low2 <- read_csv("rep1_3kx_2.csv")
low3 <- read_csv("rep1_3kx_3.csv")
low4 <- read_csv("rep1_3kx_4.csv")
low5 <- read_csv("rep2_5kx_1.csv")
low6 <- read_csv("rep2_5kx_2.csv")
low7 <- read_csv("rep2_5kx_3.csv")
low8 <- read_csv("rep2_5kx_4.csv")
low9 <- read_csv("rep3_5kx_1.csv")
low10 <- read_csv("rep3_5kx_2.csv")
low11 <- read_csv("rep3_5kx_3.csv")
reg1 <- read_csv("ctrl_5kx_1.csv")
reg2 <- read_csv("ctrl_5kx_2.csv")
reg3 <- read_csv("ctrl_5kx_3.csv")

alllow <- read_csv("All low O2 widths.csv")
allreg <- read_csv("All reg O2 widths.csv")

oxygen <- read_csv("1% vs 21% widths.csv")
```

## Data distributions

Below are histograms for each sampled image. 
```{r pressure, echo=FALSE}
hist(low1$Length)
hist(low2$Length)
hist(low3$Length)
hist(low4$Length)
hist(low5$Length)
hist(low6$Length)
hist(low7$Length)
hist(low8$Length)
hist(low9$Length)
hist(low10$Length)
hist(low11$Length)
hist(reg1$Length)
hist(reg2$Length)
hist(reg3$Length)
```

All data appear relatively normal, with some displaying moderate right skew. This indicates that most measurements are on the smaller end.

## Analysis

For convenience, I grouped all observations from each condition into one dataset. Below are summary statistics and graphs showing their distribution.

```{r}
summary(alllow)
summary(allreg)

ggplot(oxygen, aes(x = O2, y = width)) + 
  geom_boxplot() +
  geom_jitter(aes(group = O2, color = O2), alpha = 0.4) +
  geom_line(aes(group = O2, color = O2), alpha = 0.4) + 
  stat_summary(fun = "mean", size = 3, geom = "point", color="red") +
  stat_summary(fun = "mean", size = 1, geom = "line", color="red", aes(group = 1)) +
  theme(legend.position = "none") +
  ylab("Cell width (uM)") +
  xlab("O2 levels")
```

In the above boxplots, the bold red dots indicate mean cell width, with a bold red line connecting the two to display how they change with oxygen concentration. See how it isn't by much-- I'm not expecting a significant result from ANOVA, let's see if I'm right!

```{r ANOVA}
oxygen.fit <- aov(width ~ O2, data=oxygen)
autoplot(oxygen.fit)
summary(oxygen.fit)
```

The autoplot graphs show no issues with ANOVA assumptions (20-30 outliers in a dataset of 2200 is inconsequential), meaning there should be no major issues with the model. Looking at the summary output, with an F statistic of 0.143 on 1 and 2198 degrees of freedom (p-value = 0.706), I can confidently say that **there is no significant difference between mean cell width at 1% O2 and 21% O2.**
